---
title: "Visualizing Effect Sizes and Confidence Intervals with effect.size.wiz"
author: "Fabian Schwendinger & Eric Lichtenstein"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Effect Size Visualization with effect.size.wiz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The **effect.size.wiz** package provides tools for visualizing effect sizes and their 95% confidence intervals in a clear and intuitive way. It includes two main plotting functions, `es_plot_v1` and `es_plot_v2`, which generate visualizations with color-coded confidence intervals and customizable design options such as color palettes, label fonts, and more.

This vignette will guide you through how to use the functions in the package to create compelling visualizations for effect sizes and confidence intervals.

## Features of the Package

The package supports:

-   Color-coded confidence intervals based on effect size.

-   Adjustable color palettes (e.g., Viridis or grayscale).

-   Speedometer-like arrows to indicate the estimated effect size.

-   Customizable axis labels, font sizes, and export options for the plots.

## Prerequisites

You need to install the **effect.size.wiz** package to use its functionality.

## Installation

To install the package, use the following commands in R:

`install.packages("remotes")remotes::install_github("FSchwendinger/effect.size.wiz")`

After the installation is complete, load the package:

```{r load_package, message = FALSE, warning = FALSE, echo = TRUE, eval = F}

library(effect.size.wiz)
```

## Usage of `es_plot_v1`

The `es_plot_v1` function creates a plot that visualizes effect sizes and their 95% confidence intervals using a color-coded gradient.

Function Syntax:

```{r es_plot_v1, message = FALSE, warning = FALSE, echo = TRUE, eval = F}
effect.size.wiz::es_plot_v1(
  effect_size,
  ci_lower,
  ci_upper,
  palette = "viridis",
  arrow_color = "black",
  axis_title_size = 12,
  axis_label_size = 10,
  save_path = NULL
)
```

### Example Usage

Let's create a simple plot using `es_plot_v1`.

```{r es_plot_v1 example, message = FALSE, warning = FALSE, echo = TRUE, eval = F}
library(effect.size.wiz)

# Define the effect size and confidence intervals
effect_size <- 0.4
ci_lower <- 0
ci_upper <- 0.8

# Create the plot
effect.size.wiz::es_plot_v1(
  effect_size,
  ci_lower,
  ci_upper,
  palette = "viridis",
  arrow_color = "black",
  axis_title_size = 12,
  axis_label_size = 10,
  save_path = "example_plot.png"
)
```

### Explanation of Parameters:

-   `effect_size`: Numeric value indicating the effect size estimate.

-   `ci_lower` and `ci_upper`: Lower and upper bounds of the 95% confidence interval.

-   `palette`: Color palette for the plot. Supported options include `"viridis"` and `"grey"`.

-   `arrow_color`: Color for the arrow representing the effect size.

-   `axis_title_size` and `axis_label_size`: Control the font size of the axis labels and title.

-   `save_path`: Optional argument to specify the file path for saving the plot. The plot will be saved as .png with dpi = 400.

-   `reverse`: Logical flag to reverse the colour scale.

## Usage of `es_plot_v2`

The `es_plot_v2` function adds a gradient shading within the 95% confidence interval, highlighting the distance from the effect size.

Function Syntax:

```{r es_plot_v2, message = FALSE, warning = FALSE, echo = TRUE, eval = F}
effect.size.wiz::es_plot_v2(
  effect_size,
  ci_lower,
  ci_upper,
  palette = "viridis",
  arrow_color = "black",
  reverse = FALSE,
  axis_title_size = 12,
  axis_label_size = 10,
  save_path = NULL
)
```

### Example Usage of `es_plot_v2`

```{r es_plot_v2 example, message = FALSE, warning = FALSE, echo = TRUE, eval = F}
# Define effect size and confidence intervals
effect_size <- -0.3
ci_lower <- -0.6
ci_upper <- -0.1

# Create the plot
effect.size.wiz::es_plot_v2(
  effect_size,
  ci_lower,
  ci_upper,
  palette = "grey",
  arrow_color = "darkred"
)
```

## Summary

The **effect.size.wiz** package is a simple and effective way to visualize effect sizes and their confidence intervals, with customisable design options for clarity. Whether you're preparing visualisations for a presentation or publication, the package provides intuitive functions for creating high-quality plots.

```{r}
sessionInfo()
```

## Additional Resources

-   [GitHub Repository for effect.size.wiz](https://github.com/FSchwendinger/effect.size.wiz)
